---
title: "It's Basketball Season"
author: "Edie Espejo"
output: html_document
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)
```

# Part 1. Webscraping

I am going to scrape NBA data and put player data into a format that I can use to compare teams using statistical methods. I will be using the following sources to retrieve data. I do not prefer one site over the other; I simply began my analysis with Basketball Reference.

  1. Basketball Reference
    - https://www.basketball-reference.com/teams/
    - https://www.basketball-reference.com/contracts/
  2. NBA Official Site
    - http://www.nba.com/teams/
  
I am constructing the following dataframes.

  1. Basketball Reference
    a. `bf_player_urls`
      - Large dataframe
      - Encompasses different leagues (NBA, BBA, more...)
      - Has historic data, some rookies
    b. `bf_team_urls`
      - Small dataframe
      - Only active franchises
      - Found out rosters are not current under these URLs
    c. `bf_salaries`
  
  2. NBA
    a. `nba_teams`
      - Took rosters from here since they are current
    b. `nba_players`
      - Large dataframe
      - All players on each of the teams


### Source 1. Basketball Reference
First, we scrape the players page on Basketball Reference and retrieve all the URLS for each of the players listed on the site.


#### Basketball Reference Player URLS
We will construct `bf_player_urls`.

```{r, include=FALSE}
# * HOME AND PLAYERS STARTING LINKS
home <- "https://www.basketball-reference.com"
players_url <- paste0(home, "/players")


# * READ IN ALL LINKS
webpage <- read_html(players_url)
results_df <- data.frame(
  cbind(
    hover=webpage %>% html_nodes("a") %>% html_text,
    url=webpage %>% html_nodes("a") %>% html_attr("href")
    )
  )


# * PARSE OUT ONLY ABC PAGES
letter_pages <- results_df[which(as.character(results_df[,1]) %in% LETTERS),]
letter_pages[,2] <- paste0(home, letter_pages[,2])
letter_pages


# * RETRIEVE LINKS FROM EACH ABC PAGE
get_letter_links <- function(abc_df_row) {
  letter <- tolower(abc_df_row[1])
  url <- abc_df_row[2]
  
  letter_webpage <- read_html(url)
  
  this_letter_df <- data.frame(
  cbind(
    hover=letter_webpage %>% html_nodes("a") %>% html_text,
    url=letter_webpage %>% html_nodes("a") %>% html_attr("href")
  )
  )

  this_letter_df <- this_letter_df[grepl(paste0("^/players/", letter), this_letter_df[,2]),]
  this_letter_df <- this_letter_df[grepl("html", this_letter_df[,2]),]
}


# * MAKE THE DATAFRAMES FOR EACH LETTER
all_link_dfs <- apply(letter_pages, 1, get_letter_links)


# * CONSOLIDATE THE DATAFRAMES
bf_player_urls <- do.call(rbind, all_link_dfs)
bf_player_urls <- bf_player_urls[!duplicated(bf_player_urls[,2]),]
bf_player_urls[,2] <- paste0(home, as.character(bf_player_urls[,2]))
```

```{r}
# * SHOW A SAMPLE
bf_player_urls[as.integer(runif(5, 0, nrow(bf_player_urls))),]
```

#### Basketball Reference Team URLS
We will construct `bf_team_urls`.

```{r, include=FALSE}
# * HOME AND FRANCHISE STARTING LINKS
franchise_url <- "https://www.basketball-reference.com/teams/"
webpage <- read_html(franchise_url)


# * ACTIVE FRANCHISES ONLY
tables <- webpage %>% html_nodes("table") %>% html_table()
active_df <- tables[[1]]
active_df <- active_df %>% filter(Lg=="NBA")
active_df <- active_df %>% filter(To=="2019")
active_df <- active_df[!duplicated(active_df$Franchise),]


# * GET LINKS OF ACTIVE FRANCHISES
page_links <- cbind(webpage %>% html_nodes("a") %>% html_text,
                    webpage %>% html_nodes("a") %>% html_attr("href")
                    )

bf_team_urls <- page_links[grepl("teams/[A-z]{3}/$", page_links[,2]),]
bf_team_urls <- data.frame(bf_team_urls)
names(bf_team_urls) <- c("team", "url")

bf_team_urls <- bf_team_urls %>% filter(team %in% active_df[,1])
bf_team_urls <- bf_team_urls[!duplicated(bf_team_urls[,1]),]

bf_team_urls$url <- paste0("https://www.basketball-reference.com", bf_team_urls[,2], "players.html")
bf_team_urls
```

```{r}
# * SHOW A SAMPLE
bf_team_urls[as.integer(runif(5, 0, nrow(bf_team_urls))),]
```

#### Basketball Reference Salaries
Continue later.
```{r}
# * HOME AND FRANCHISE STARTING LINKS
payroll <- "https://www.basketball-reference.com/contracts/"
webpage <- read_html(payroll)


# * TOP SALARY TABLE
tables <- webpage %>% html_nodes("table") %>% html_table()

# * PER TEAM
team_list <- webpage %>% html_nodes("table") %>% html_nodes("a")
team_payroll_urls <- cbind(
  team=team_list %>% html_text(),
  url=team_list %>% html_attr("href")
  )
team_payroll_urls <- data.frame(team_payroll_urls)
team_payroll_urls[,2] <- paste0(home, team_payroll_urls[,2])

generate_salaries <- function(payroll_row) {
  this_team <- as.character(payroll_row[1])
  this_page <- as.character(payroll_row[2])
  webpage <- read_html(this_page)

  salary_table <- data.frame(webpage %>% html_nodes("table") %>% html_table())
  names(salary_table) <- salary_table[1,]
  salary_table <- salary_table[-1,]
  salary_table <- salary_table %>% mutate(team=this_team)
  
  salary_table
}

team_payrolls <- apply(team_payroll_urls, 1, generate_salaries)
team_payrolls <- do.call(rbind, team_payrolls)
team_payrolls <- team_payrolls[-which(team_payrolls$Player==""),]
team_payrolls <- team_payrolls[-which(team_payrolls$Player=="Team Totals"),]
```


```{r}
western_conf <- team_payrolls %>% 
  filter(team %in% c("Golden State Warriors", "Dallas Mavericks", "Denver Nuggets", "Houston Rockets", "Memphis Grizzlies", "Minnesota Timberwolves", "New Orleans Pelicans", "Portland Trail Blazers", "San Antonio Spurs", "Utah Jazz", "Los Angeles Lakers", "Los Angeles Clippers", "Sacramento Kings", "Phoenix Suns", "Oklahoma City Thunder")) %>%
  select(Player, millions, team) %>%
  arrange(-millions) %>%
  filter(!is.na(millions))
write.csv(western_conf, "../../../../GitHub/ph142_personal/midterm_2/western_nba.csv")
```

```{r}
# * SHOW A SAMPLE
team_payrolls[as.integer(runif(5, 0, nrow(team_payrolls))),]
```

```{r}
library(ggplot2)
names(team_payrolls)[3] = "Current"

team_payrolls[3] <- as.numeric(gsub("\\$|\\,", "", as.vector(unlist(team_payrolls[3]))))
team_payrolls <- team_payrolls %>% mutate(millions=Current/1000000)

ggplot(team_payrolls, aes(x=millions)) +
  geom_histogram(bins=30, na.rm=TRUE) +
  xlab("millions of $") +
  ylab("frequency") +
  ggtitle("How much are NBA players making this year?") +
  facet_wrap(~team)
```

```{r}
team_payrolls %>% arrange(-Current)
```

### Source 2. NBA

#### NBA Teams

```{r, include=FALSE}
# * SPECIFY STARTING PAGE
nba_teams <- "http://www.nba.com/teams"
webpage <- read_html(nba_teams)

# * SCRAPE THE COLUMN OF TEAMS
teams <- webpage %>% html_nodes("div.team__list") %>% html_nodes("a") %>% html_text()
team_urls <- webpage %>% html_nodes("div.team__list") %>% html_nodes("a") %>% html_attr("href")

# * CREATE DATAFRAME
team_df <- data.frame(cbind(teams, team_urls))
team_df$team_urls <- paste0("http://www.nba.com", team_df$team_urls)
```

```{r}
# * SHOW A SAMPLE
team_df[as.integer(runif(5, 0, nrow(team_df))),]
```

#### NBA Players
```{r, include=FALSE}
retrieve_players <- function(team_row) {
  this_team <- as.character(team_row[,1])
  webpage <- read_html(as.character(team_row[,2]))
  
  team_roster <- cbind(player=webpage %>% html_nodes("a") %>% html_attr("title"),
                     url=webpage %>% html_nodes("a") %>% html_attr("href")
                     )
  team_roster <- team_roster[!is.na(team_roster[,1]),]
  team_roster <- team_roster[grepl("players", team_roster[,2]),]
  
  team_roster <- cbind(team_roster, team=rep(this_team, nrow(team_roster)))
  
  team_roster
}

retrieve_players(team_df[grepl("Warriors", team_df[,1]),])

nba_players <- do.call(rbind, lapply(1:nrow(team_df), function(x) retrieve_players(team_df[x,])))
nba_players <- data.frame(nba_players)
```

```{r}
# * SHOW A SAMPLE
nba_players[as.integer(runif(5, 0, nrow(nba_players))),]
```

# Part 2. Warriors
First, I'm going to put the links all together for each of the players. Not all of the players will have Basketball Reference links.

```{r, include=FALSE}
# * NBA WARRIORS DATA
current_dubs <- nba_players %>% filter(grepl("Warriors", team)) %>% rename(nba_url=url) %>% mutate(nba_url=paste0("http://www.nba.com", nba_url))
current_dubs <- current_dubs[!duplicated(current_dubs$player),]

# * BASKETBALL REFERENCE WARRIORS DATA
bball_ref <- bf_player_urls %>% filter(hover %in% current_dubs$player) %>% rename(bball_ref_url=url, player=hover)

# * CONSOLIDATE
joint_df <- full_join(current_dubs, bball_ref, by="player")
joint_df
```

```{r}
```

# Appendix

#### Helpful Links
```{r}
payroll <- "https://www.basketball-reference.com/contracts/"
nba_teams <- "http://www.nba.com/teams"
league_index <- "https://www.basketball-reference.com/leagues/"
dubs <- "https://www.basketball-reference.com/teams/GSW/players.html"
```

